---
title: "Troubleshooting Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Troubleshooting Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This guide helps you diagnose and resolve common issues when working with putior. Each section covers a specific category of problems with practical solutions.

## Contents

- [Most Common Issues](#most-common-issues) - Start here!
- [Quick Diagnostics](#quick-diagnostics)
- [Annotation Syntax Errors](#annotation-syntax-errors)
- [File Pattern Matching Issues](#file-pattern-matching-issues)
- [Diagram Rendering Problems](#diagram-rendering-problems)
- [Multi-Language Support Issues](#multi-language-support-issues)
- [Performance Tips](#performance-tips-for-large-codebases)
- [Debugging with Logging](#debugging-with-logging)
- [FAQ](#frequently-asked-questions)
- [Getting Help](#getting-help)

---

## Most Common Issues

These are the top 5 issues users encounter. Check these first!

### 1. Annotations Not Detected

**Frequency:** 🔴 Very Common

**Problem:** `put()` returns an empty data frame even though you have annotations in your files.

**Cause:** Wrong comment prefix for your file type.

**Solution:** Use the correct comment prefix for your language:

| Language | Correct Prefix | Example |
|----------|---------------|---------|
| R, Python | `# put` | `# put label:"Step 1"` |
| SQL, Lua | `-- put` | `-- put label:"Step 1"` |
| JavaScript, Go, C | `// put` | `// put label:"Step 1"` |
| MATLAB | `% put` | `% put label:"Step 1"` |

```{r}
# Check what prefix your file type needs
get_comment_prefix("sql")  # Returns "--"
get_comment_prefix("js")   # Returns "//"
```

---

### 2. Nodes Not Connecting in Diagram

**Frequency:** 🔴 Very Common

**Problem:** Diagram shows isolated nodes instead of a connected workflow.

**Cause:** Typo in input/output names - they must match exactly.

**Solution:** Check that output of one node matches input of the next:

```r
# WRONG - Names don't match
# put id:"step1", output:"data.csv"      # Outputs "data.csv"
# put id:"step2", input:"Data.csv"       # Expects "Data.csv" (case mismatch!)

# CORRECT - Exact match
# put id:"step1", output:"data.csv"
# put id:"step2", input:"data.csv"
```

**Diagnostic:**
```{r}
workflow <- put("./src/")
# Check for mismatches
unique(workflow$output)
unique(workflow$input)
```

---

### 3. Diagram Not Rendering

**Frequency:** 🟡 Occasional

**Problem:** Mermaid code appears as text instead of a diagram.

**Cause:** Mermaid.js not loaded, or syntax error in generated code.

**Solution:**

1. **For pkgdown sites:** Ensure Mermaid is included in `_pkgdown.yml`
2. **For R Markdown:** Use raw output and a Mermaid-enabled viewer
3. **Quick test:** Paste the output into [mermaid.live](https://mermaid.live)

```{r}
# Get raw Mermaid code to debug
mermaid_code <- put_diagram(workflow, output = "raw")
cat(mermaid_code)  # Copy this to mermaid.live to test
```

---

### 4. Auto-Detection Missing Expected Patterns

**Frequency:** 🟡 Occasional

**Problem:** `put_auto()` doesn't detect file reads/writes you expected.

**Cause:** Library or function not in detection patterns.

**Solution:**

1. Check if your library is supported:
```{r}
patterns <- get_detection_patterns("r")
# Search for your function
grep("readr", sapply(patterns$input, `[[`, "func"), value = TRUE)
```

2. Use manual annotations for unsupported patterns
3. Request new patterns via [GitHub issue](https://github.com/pjt222/putior/issues)

---

### 5. Wrong Language Detection

**Frequency:** 🟢 Rare

**Problem:** File parsed with wrong comment syntax.

**Cause:** Unusual file extension or ambiguous extension (e.g., `.m` for MATLAB vs Objective-C).

**Solution:**

```{r}
# Check what language is detected for your extension
ext_to_language("m")   # Returns "matlab"
ext_to_language("jl")  # Returns "julia"

# List all supported extensions
get_supported_extensions()
```

If your extension isn't recognized, use the standard extension or add manual annotations with the correct comment prefix.

---

## Quick Diagnostics

Before diving into specific issues, run these quick checks:

```{r}
library(putior)

# 1. Check if annotations are found
workflow <- put("./src/", validate = TRUE, include_line_numbers = TRUE)
nrow(workflow)  # Should be > 0 if annotations exist

# 2. Test a single annotation
is_valid_put_annotation('# put id:"test", label:"Test Node"')  # Should be TRUE

# 3. Check supported extensions
get_supported_extensions()

# 4. Enable detailed logging
set_putior_log_level("DEBUG")
workflow <- put("./src/")  # Now shows detailed output
```

---

## Annotation Syntax Errors

### Missing or Mismatched Quotes

**Problem**: Annotation is not recognized or properties are missing.

```r
# WRONG - No quotes around values
# put id:my_node, label:My Process

# WRONG - Mismatched quotes
# put id:"my_node', label:"My Process"

# WRONG - Missing closing quote
# put id:"my_node, label:"My Process"
```

**Solution**: Always use matching quotes around all values.

```r
# CORRECT - Double quotes
# put id:"my_node", label:"My Process"

# CORRECT - Single quotes
# put id:'my_node', label:'My Process'

# CORRECT - Mixed (but consistent per value)
# put id:"my_node", label:'My Process'
```

**Diagnostic**:

```{r}
# Test your annotation syntax
is_valid_put_annotation('# put id:"my_node", label:"My Process"')  # TRUE
is_valid_put_annotation('# put id:my_node, label:My Process')      # FALSE
```

### Invalid Node Types

**Problem**: Validation warning about unusual node_type.

```r
# Triggers warning - non-standard type
# put id:"step1", label:"Process", node_type:"transform"
```

**Solution**: Use standard node types for consistency.

```r
# Standard node types
# put id:"step1", label:"Load Data", node_type:"input"
# put id:"step2", label:"Transform", node_type:"process"
# put id:"step3", label:"Save Results", node_type:"output"
# put id:"step4", label:"Check Quality", node_type:"decision"
# put id:"step5", label:"Start Pipeline", node_type:"start"
# put id:"step6", label:"End Pipeline", node_type:"end"
```

**Note**: Custom node types work but generate warnings. Suppress with `validate = FALSE`:

```{r}
workflow <- put("./src/", validate = FALSE)
```

### Commas Inside Values

**Problem**: Values containing commas are truncated.

```r
# WRONG - Comma breaks the label
# put id:"step1", label:Load, clean, transform data
```

**Solution**: Enclose values with commas in quotes.

```r
# CORRECT - Quotes protect the comma
# put id:"step1", label:"Load, clean, transform data"
```

### Empty or Missing IDs

**Problem**: Warning about empty ID.

```r
# WRONG - Empty ID string
# put id:"", label:"My Process"

# OK - Missing ID (auto-generated UUID)
# put label:"My Process"
```

**Solution**: Either omit the ID entirely (auto-generated) or provide a valid one.

```{r}
# With auto-generated ID
workflow <- put("./src/")
print(workflow$id)  # Shows UUIDs like "a1b2c3d4-..."

# With explicit IDs
# put id:"step1", label:"First Step"
# put id:"step2", label:"Second Step"
```

### Duplicate IDs

**Problem**: Warning about duplicate node IDs.

```r
# In file1.R
# put id:"load_data", label:"Load from CSV"

# In file2.R
# put id:"load_data", label:"Load from JSON"  # DUPLICATE!
```

**Solution**: Use unique IDs across all files in your workflow.

```r
# In file1.R
# put id:"load_csv", label:"Load from CSV"

# In file2.R
# put id:"load_json", label:"Load from JSON"
```

### Multiline Annotation Issues

**Problem**: Multiline annotations not being parsed correctly.

```r
# WRONG - Missing backslash
# put id:"step1", label:"Process",
#     input:"data.csv", output:"results.csv"

# WRONG - Backslash not at end of line
# put id:"step1", label:"Process", \
#     input:"data.csv"

# WRONG - Missing comment prefix on continuation
# put id:"step1", label:"Process", \
    input:"data.csv"
```

**Solution**: Use proper multiline syntax.

```r
# CORRECT - Backslash at end, comment prefix on each line
# put id:"step1", label:"Process Data", \
#     input:"data.csv,config.json", \
#     output:"results.csv"
```

**For SQL files**:

```sql
-- put id:"etl_job", label:"ETL Process", \
--     input:"source_table", \
--     output:"target_table"
```

**For JavaScript files**:

```javascript
// put id:"handler", label:"API Handler", \
//     input:"request.json", \
//     output:"response.json"
```

---

## File Pattern Matching Issues

### Files Not Found

**Problem**: `put()` returns empty result or "no files found" warning.

```{r}
workflow <- put("./src/")
#> Warning: No files matching pattern '\.(R|r|py|sql|sh|jl)$' found in: ./src/
```

**Diagnostic Steps**:

```{r}
# 1. Check if directory exists
dir.exists("./src/")

# 2. List files in directory
list.files("./src/", recursive = TRUE)

# 3. Check what patterns match
list.files("./src/", pattern = "\\.(R|r|py)$")
```

**Solutions**:

```{r}
# Use correct path (absolute or relative)
workflow <- put("/full/path/to/src/")

# Scan recursively for subdirectories
workflow <- put("./src/", recursive = TRUE)

# Check if single file path works
workflow <- put("./src/script.R")
```

### Wrong File Extensions

**Problem**: Your files use extensions not in the default pattern.

```{r}
# Default pattern only matches: R, r, py, sql, sh, jl
workflow <- put("./src/")  # Won't find .js, .ts, .go files
```

**Solution**: Specify a custom pattern for your file types.

```{r}
# JavaScript/TypeScript files
workflow <- put("./src/", pattern = "\\.(js|ts|jsx|tsx)$")

# Go files
workflow <- put("./src/", pattern = "\\.go$")

# Rust files
workflow <- put("./src/", pattern = "\\.rs$")

# Multiple languages
workflow <- put("./src/", pattern = "\\.(R|py|js|ts|go|rs)$")

# All supported extensions
all_exts <- paste(get_supported_extensions(), collapse = "|")
pattern <- paste0("\\.(", all_exts, ")$")
workflow <- put("./src/", pattern = pattern)
```

### Case Sensitivity

**Problem**: Files not found due to case mismatch.

```{r}
# Pattern is case-sensitive by default
list.files("./src/", pattern = "\\.R$")   # Finds script.R
list.files("./src/", pattern = "\\.r$")   # Finds script.r
```

**Solution**: Include both cases in pattern.

```{r}
# Match both .R and .r
workflow <- put("./src/", pattern = "\\.(R|r)$")

# Or use the default pattern which includes both
workflow <- put("./src/")  # Default: "\\.(R|r|py|sql|sh|jl)$"
```

---

## Diagram Rendering Problems

### Mermaid Syntax Errors

**Problem**: Diagram doesn't render or shows syntax error.

**Diagnostic**:

```{r}
# Get raw Mermaid code to inspect
workflow <- put("./src/")
mermaid_code <- put_diagram(workflow, output = "raw")
cat(mermaid_code)
```

**Common Causes**:

1. **Special characters in labels**: Mermaid has reserved characters.

```{r}
# Problem: Quotes in labels
# put id:"step1", label:"Load "raw" data"

# Solution: Use single quotes or escape
# put id:"step1", label:"Load 'raw' data"
# put id:"step1", label:"Load raw data"
```

2. **Reserved words in IDs**: Some words are reserved in Mermaid.

```{r}
# Problem: Reserved word as ID
# put id:"end", label:"Finish"
# put id:"class", label:"Classify"

# Solution: Use different IDs
# put id:"end_step", label:"Finish"
# put id:"classify", label:"Classify"
```

### Theme Issues

**Problem**: Diagram colors don't appear or look wrong.

```{r}
# Check available themes
get_diagram_themes()
#> $light, $dark, $auto, $minimal, $github
#> $viridis, $magma, $plasma, $cividis (colorblind-safe)
```

**Solutions**:

```{r}
# Try a different theme
put_diagram(workflow, theme = "github")
put_diagram(workflow, theme = "minimal")

# Disable node styling for plain diagrams
put_diagram(workflow, style_nodes = FALSE)

# Check if theme is compatible with your viewer
# "github" works best on GitHub README
# "dark" for dark mode applications
# "minimal" for printing/PDFs
# "viridis", "magma", "plasma", "cividis" for colorblind accessibility
```

### Missing Connections

**Problem**: Nodes appear but arrows don't connect them.

**Cause**: Input/output values don't match between nodes.

```{r}
# Problem: Typo in file name
# put id:"step1", output:"data.csv"
# put id:"step2", input:"Data.csv"  # Case mismatch!

# Solution: Match exactly
# put id:"step1", output:"data.csv"
# put id:"step2", input:"data.csv"
```

**Diagnostic**:

```{r}
# Check what inputs and outputs are defined
workflow <- put("./src/")
print(workflow[, c("id", "input", "output")])
```

### Artifacts Not Showing

**Problem**: Data files (artifacts) don't appear in diagram.

```{r}
# Enable artifacts display
put_diagram(workflow, show_artifacts = TRUE)

# Artifacts are intermediate files that connect nodes
# They're hidden by default for cleaner diagrams
```

---

## Multi-Language Support Issues

### Wrong Comment Prefix Detected

**Problem**: Annotations not found in SQL, JavaScript, or other files.

**Cause**: Using wrong comment syntax for the language.

```sql
-- WRONG for SQL - using hash instead of dash
# put id:"query", label:"Run Query"

-- CORRECT for SQL
-- put id:"query", label:"Run Query"
```

```javascript
// WRONG for JavaScript - using hash
# put id:"handler", label:"Handle Request"

// CORRECT for JavaScript
// put id:"handler", label:"Handle Request"
```

**Check the correct prefix**:

```{r}
# Get comment prefix for any extension
get_comment_prefix("sql")   # "--"
get_comment_prefix("js")    # "//"
get_comment_prefix("m")     # "%"
get_comment_prefix("py")    # "#"
get_comment_prefix("go")    # "//"
get_comment_prefix("rs")    # "//"
```

### Reference Table

| Extension | Comment Prefix | Example |
|-----------|---------------|---------|
| `.R`, `.r` | `#` | `# put id:"step", label:"Process"` |
| `.py` | `#` | `# put id:"step", label:"Process"` |
| `.sql` | `--` | `-- put id:"query", label:"Query"` |
| `.js`, `.ts` | `//` | `// put id:"func", label:"Function"` |
| `.go` | `//` | `// put id:"main", label:"Main"` |
| `.rs` | `//` | `// put id:"fn", label:"Handler"` |
| `.m` (MATLAB) | `%` | `% put id:"calc", label:"Compute"` |
| `.tex` | `%` | `% put id:"sec", label:"Section"` |
| `.sh` | `#` | `# put id:"cmd", label:"Command"` |
| `.lua` | `--` | `-- put id:"fn", label:"Function"` |

### Unsupported Extension

**Problem**: File type not automatically recognized.

```{r}
# Check if extension is supported
ext <- "xyz"
get_comment_prefix(ext)  # Returns "#" as fallback

# List all supported extensions
get_supported_extensions()
```

**Escape Hatches for Unsupported File Types:**

**Option 1: Use fallback `#` prefix** (if your language supports `#` comments)
```
# In your .xyz file (if it supports # comments)
# put id:"step1", label:"Process Data"
```

**Option 2: Create a wrapper annotation file**
```r
# workflow-annotations.R
# This file documents the workflow for unsupported file types

# put id:"xyz_step1", label:"Process in XYZ format", input:"data.xyz", output:"result.xyz"
# put id:"xyz_step2", label:"Convert output", input:"result.xyz", output:"final.csv"

# Then scan this file instead
workflow <- put("workflow-annotations.R")
```

**Option 3: Use text input for ad-hoc annotations**
```{r}
# Define workflow without files
workflow <- put(text = '
# put id:"step1", label:"Custom Step 1", output:"data.xyz"
# put id:"step2", label:"Custom Step 2", input:"data.xyz"
')
put_diagram(workflow)
```

**Option 4: Request new language support**

[Open a GitHub issue](https://github.com/pjt222/putior/issues) with:
- Language name and file extension
- Comment syntax used by the language
- Example code showing typical file I/O patterns (for auto-detection)

---

## Performance Tips for Large Codebases

### Slow Scanning

**Problem**: `put()` takes too long on large projects.

**Solutions**:

```{r}
# 1. Limit to specific directories
workflow <- put("./src/core/")  # Instead of ./

# 2. Use specific file patterns
workflow <- put("./src/", pattern = "\\.R$")  # Only R files

# 3. Limit to top-level directory if subdirs not needed
workflow <- put("./src/", recursive = FALSE)

# 4. Disable validation for faster scanning
workflow <- put("./src/", validate = FALSE)
```

### Memory Issues with Large Workflows

**Problem**: Many nodes cause memory or rendering issues.

**Solutions**:

```{r}
# 1. Split into smaller subgraphs
workflow_etl <- put("./src/etl/")
workflow_ml <- put("./src/ml/")

# Generate separate diagrams
put_diagram(workflow_etl, file = "etl_workflow.md")
put_diagram(workflow_ml, file = "ml_workflow.md")

# 2. Hide artifacts for cleaner diagrams
put_diagram(workflow, show_artifacts = FALSE)

# 3. Use simpler theme
put_diagram(workflow, theme = "minimal", style_nodes = FALSE)
```

### Reducing Annotation Overhead

**Best Practices**:

```{r}
# Annotate key steps only, not every function
# Good: Major pipeline stages
# put id:"ingest", label:"Data Ingestion", node_type:"input"
# put id:"transform", label:"Data Transformation", node_type:"process"
# put id:"export", label:"Export Results", node_type:"output"

# Avoid: Every small helper function
# (Don't annotate utility functions unless they're significant)
```

---

## Debugging with Logging

### Enable Detailed Logging

```{r}
# Install logger package (optional but recommended)
install.packages("logger")

# Set log level
set_putior_log_level("DEBUG")  # Maximum detail
set_putior_log_level("INFO")   # Progress updates
set_putior_log_level("WARN")   # Issues only (default)
set_putior_log_level("ERROR")  # Fatal errors only
```

### Per-Call Logging

```{r}
# Override log level for a single call
workflow <- put("./src/", log_level = "DEBUG")
put_diagram(workflow, log_level = "INFO")
```

### What Each Level Shows

| Level | Information Shown |
|-------|-------------------|
| `DEBUG` | File-by-file processing, pattern matching, parsing steps |
| `INFO` | Scan started, files found, nodes extracted, diagram generated |
| `WARN` | Validation issues, missing dependencies |
| `ERROR` | Fatal errors that stop execution |

### Debugging Example

```{r}
# Problem: Annotations not being found

# 1. Enable debug logging
set_putior_log_level("DEBUG")

# 2. Run the scan
workflow <- put("./problem_file.R", include_line_numbers = TRUE)

# 3. Check the logs for:
#    - "No PUT annotations found in..." (wrong syntax?)
#    - "Found X PUT annotation(s)..." (they exist!)
#    - "Processing file:..." (file is being scanned)

# 4. Reset to normal
set_putior_log_level("WARN")
```

### Logger Not Installed

If you don't have the `logger` package installed, putior works silently without logging. Install it for debugging:

```{r}
install.packages("logger")

# Verify it's working
library(logger)
set_putior_log_level("INFO")
workflow <- put("./src/")
#> INFO [2024-01-15 10:30:00] Starting PUT annotation scan
#> INFO [2024-01-15 10:30:00] Found 3 file(s) to scan
#> INFO [2024-01-15 10:30:01] Scan complete: found 5 workflow node(s)
```

---

## Frequently Asked Questions

### Q: Why are my annotations not found?

**A**: Check these common causes:

1. **Wrong comment prefix** for file type (see [Multi-Language Support](#multi-language-support-issues))
2. **Missing quotes** around property values
3. **File not matching pattern** (check `list.files()`)
4. **Typo in "put"** keyword (must be lowercase: `# put` not `# PUT`)

```{r}
# Validate syntax
is_valid_put_annotation('# put id:"test", label:"Test"')  # TRUE
is_valid_put_annotation('# PUT id:"test", label:"Test"')  # FALSE (uppercase)
```

### Q: How do I connect nodes between files?

**A**: Use matching input/output file names:

```r
# In 01_load.R
# put id:"load", label:"Load Data", output:"raw_data.csv"

# In 02_process.R
# put id:"process", label:"Process", input:"raw_data.csv", output:"clean_data.csv"
```

### Q: What if I don't specify an output?

**A**: putior automatically uses the file name as the output.

```r
# In process_data.R
# put id:"process", label:"Process Data", input:"raw.csv"
# Output defaults to "process_data.R"

# In analyze.R
# put id:"analyze", label:"Analyze", input:"process_data.R"
# This creates a connection from process_data.R to analyze.R
```

### Q: Can I use putior in CI/CD?

**A**: Yes, putior works in headless environments:

```{r}
# Generate diagram file for documentation
workflow <- put("./src/", recursive = TRUE)
put_diagram(workflow,
            output = "file",
            file = "docs/workflow.md",
            theme = "github")
```

### Q: How do I track in-memory variables?

**A**: Use the `.internal` extension:

```r
# Variables created during script execution
# put id:"create", output:"dataset.internal, dataset.RData"
dataset <- data.frame(x = 1:100)
save(dataset, file = "dataset.RData")

# Next script uses the saved file, not .internal
# put id:"analyze", input:"dataset.RData", output:"results.internal"
load("dataset.RData")
results <- summary(dataset)
```

**Important**: `.internal` files can only be outputs, never inputs between scripts.

### Q: Why does the diagram look different on GitHub vs locally?

**A**: Different Mermaid renderers interpret themes differently.

```{r}
# For GitHub README
put_diagram(workflow, theme = "github")

# For local HTML
put_diagram(workflow, theme = "light")

# For universal compatibility
put_diagram(workflow, theme = "minimal", style_nodes = FALSE)
```

### Q: Can I use custom properties?

**A**: Yes, putior preserves any properties you define:

```r
# put id:"train", label:"Train Model", \
#     node_type:"process", \
#     runtime:"2h", \
#     gpu:"required", \
#     priority:"high"
```

```{r}
workflow <- put("./src/")
print(workflow$runtime)   # "2h"
print(workflow$gpu)       # "required"
print(workflow$priority)  # "high"
```

### Q: How do I suppress validation warnings?

**A**: Disable validation:

```{r}
# Suppress all validation warnings
workflow <- put("./src/", validate = FALSE)

# Or suppress specific warnings with suppressWarnings()
workflow <- suppressWarnings(put("./src/"))
```

---

## Getting Help

If you're still stuck:

1. **Check the vignettes**:
   - [Quick Start](quick-start.html) - First diagram in 2 minutes
   - [Annotation Guide](annotation-guide.html) - Complete syntax reference
   - [Quick Reference](quick-reference.html) - Cheat sheet
   - [Features Tour](features-tour.html) - Advanced features
   - [API Reference](api-reference.html) - Complete function docs

2. **Use the sandbox**:
   ```{r}
   run_sandbox()  # Interactive experimentation
   ```

3. **Enable debug logging** and check output carefully

4. **Report issues** on GitHub with:
   - R version (`R.version.string`)
   - putior version (`packageVersion("putior")`)
   - Minimal reproducible example
   - Debug log output

```{r}
# Gather diagnostic info
sessionInfo()
packageVersion("putior")
get_supported_extensions()
```
